Automatic discovery of cell types and microcircuitry from neural connectomics

### 2014-12-10 docs/paper.tex
updates to experiment runners

\begin{abstract}
  Neural connectomics has begun producing massive amounts of data,
  necessitating new analysis methods to discover the biological and
  computational structure. It has long been assumed that discovering
  neuron types and their relation to microcircuitry is crucial to
  understanding neural function. Here we developed a nonparametric
  Bayesian technique that identifies neuron types and microcircuitry
  patterns in connectomics data. It combines the information
  traditionally used by biologists, including connectivity, cell body
  location and the spatial distribution of synapses, in a
  principled and probabilistically-coherent manner. We show that the
  approach recovers known neuron types in the retina and enables
  predictions of connectivity, better than simpler algorithms. It also
  can reveal interesting structure in the nervous system of
  C. elegans, and automatically discovers the structure of a
  microprocessor.  Our approach extracts structural meaning from
  connectomics, enabling new approaches of automatically deriving
  anatomical insights from these emerging datasets.
\end{abstract}

@@

We primarily focus on the recently-released mouse retina connectome
\autocite{Helmstaedter2013}, but additionally examine the C. elegans
connectome \autocite{White1986}, and then ``connectome'' of a
classical microprocessor \autocite{James2010}. Comparing the cell
types discovered by the algorithms with those obtained manually by
human anatomists reveals a high degree of agreement. We thus present a
scalable probabilistic approach to infer microcircuitry from
connectomics data available today and in the future.

@@

\subsection*{Types and connectivity in artificial structures}
\begin{figure}
  \centering 
  \centerline{\includegraphics[width=4in]{mos6502.ai}}
  \caption{Discovering connectivity and type the 6502 microprocessor.
    a.) is the micrograph of the original microprocessor, with the
    region containing the registers under study highlighted. b.) Our
    graph consists of the interconnections of MOS field-effect
    transistors with three terminals, Gate, C1, and C2. Reconstruction
    technique did not permit resolution of C1 and C2 into source and
    drain. c.) The spatial distribution of the transistors in each
    cluster show a clear pattern d.) The clusters and connectivity
    versus distance for connections between Gate and C1, Gate and C2,
    and C1 and C2 terminals on a transistor. Red and teal types have a
    terminal pulled down to ground and mostly function as
    inverters. Purple class are clocked, stateful transistors, orange
    control the ALU and yellow control the special data bus (SDB).}
  \label{fig:mos6502}
\end{figure}

To show the applicability of our method to other connectome-style
datasets, we obtained the spatial location and interconnectivity of
the transistors in a classic microprocessor, the MOS Technology 6502
(used in the Apple II) \autocite{James2010}. Computer architects use
common patterns of transistors when designing circuits, with each
transistor having a ``type'' in the circuit. We identified a region of
the processor with complex but known structure containing the primary
8-bit registers X, Y, and S (fig~\ref{fig:mos6502}).

Our algorithm identifies areas of spatial homogeneity that mirror the
known structure in the underlying architectural circuit, segmenting
transistor types recognizable to computer architects. Using the
original schematics, we see that one identified type contains the
``clocked'' transistors, which retain digital state. Two other types
contain transistors with pins C1 or C2 connected to ground, mostly
serving as inverters.  An additional identified type controls the
behavior of the three registers of interest (X, Y, and S) with respect
to the SB data bus, either allowing them to latch or drive data from
the bus. The repeat patterns of spatial connectivity are visible in
figure~\ref{fig:mos6502}c, showing the man-made horizontal and
vertical layout of the same types of transistors.

>>>>

\begin{abstract}
  Neural connectomics has begun producing massive amounts of data,
  necessitating new analysis methods to discover the biological and
  computational structure. It has long been assumed that discovering
  neuron types and their relation to microcircuitry is crucial to
  understanding neural function. Here we developed a nonparametric
  Bayesian technique that identifies neuron types and microcircuitry
  patterns in connectomics data. It combines the information
  traditionally used by biologists, including connectivity, cell body
  location and the spatial distribution of synapses, in a
  principled and probabilistically-coherent manner. We show that the
  approach recovers known neuron types in the retina and enables
  predictions of connectivity, better than simpler algorithms. It also
  can reveal interesting structure in the nervous system of
  C. elegans.
  Our approach extracts structural meaning from
  connectomics, enabling new approaches of automatically deriving
  anatomical insights from these emerging datasets.
\end{abstract}

@@

We primarily focus on the recently-released mouse retina connectome
\autocite{Helmstaedter2013}, but additionally examine the C. elegans
connectome \autocite{White1986}. 
Comparing the cell
types discovered by the algorithms with those obtained manually by
human anatomists reveals a high degree of agreement. We thus present a
scalable probabilistic approach to infer microcircuitry from
connectomics data available today and in the future.

@@
